---
title: "IPUMS 2022 ACS Data Analysis Using Ratio Estimation"
subtitle: "Estimating State-Level Populations Using Ratio Estimation"
author: 
  - Yizhe Chen
  - Charlie Zhang
  - Qizhou Xie
thanks: "Code and data are available at: [https://github.com/YizheChenUT/Doctoral-Degree-Analysis.git](https://github.com/YizheChenUT/Doctoral-Degree-Analysis.git)"
date: today
date-format: long
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

```

# Introduction
This analysis explores the use of the ratio estimator method through R language [@citeR] to evaluate population characteristics based on the 2022 American Community Survey (ACS) data obtained from IPUMS USA [@ipumsusa]. By using detailed state-level data on doctoral degree holders and total respondents, the study aims to estimate population metrics for states across the U.S. The method assumes that the relationship observed in a reference state (California) can be extended to other states to derive meaningful approximations.

The dataset includes variables such as STATEICP (state identifiers) and EDUCD (educational attainment), enabling the calculation of a ratio for doctoral degree holders relative to the total population in California. This ratio is applied to other states, allowing for estimations of their total respondent numbers.

The analysis also highlights discrepancies between estimated and actual values, investigating contributing factors such as demographic diversity, sampling variability, and state-level educational trends. This study provides a foundation for understanding the application of ratio estimators in demographic analysis and underscores the potential challenges when generalizing from a single-state reference point.

# Instructions on How to Obtain the Data
To obtain the IPUMS 2022 ACS data [@ipumsusa]:

- 1. Visit the [IPUMS USA website](https://usa.ipums.org/usa/).

- 2. Register for an account or log in if you already have one.

- 3. Select the 2022 ACS dataset and choose relevant variables such as:
   - `STATEICP`: State codes for each respondent.
   - `EDUCD`: Educational attainment with detailed categories, including doctoral degrees.

- 4. Download the dataset in CSV format. The file may be compressed, so you need to unzip it using:
   `gunzip usa_00003.csv.gz`

# Overview of the Ratio Estimators Approach
The ratio estimator approach is used when we want to estimate unknown totals in a population based on the known total in a sample and some measured characteristics.

We know the total number of respondents in California is 391,171. We can calculate the ratio of doctoral degree holders in California to the total respondents, then apply this ratio to estimate the total number of respondents in other states.

Ratio = Number of doctoral holders in California / 391171

We apply this ratio to the number of doctoral degree holders in other states to estimate their total respondent numbers.

# Estimates and the Actual Number of Respondents
We computed the estimated total respondents based on the ratio estimator approach. @tbl-estimated-actual-doctoral showing the estimated number of respondents and the actual number of doctoral degree holders in each state.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-estimated-actual-doctoral
#| tbl-cap: "Estimated number of respondents and the actual number of doctoral degree holders in each state"
 
# Load the comparison results and display the table
comparison <- read_csv(here::here("comparison_results.csv")) %>%
  select(-estimated_doctoral) %>%
  rename(
    `State` = STATEICP,
    `Estimated Total Respondents` = estimate_total,
    `Doctoral Degree Holders` = count_doctoral,
    `Difference` = difference
  )

knitr::kable(comparison, format = "pipe", align = "l")
```
 
# Explanation of Differences

There are several reasons why the estimated total number of respondents may differ from the actual numbers:

- State-level variations: The proportion of doctoral degree holders in California may not be representative of the proportion in other states. States with a higher or lower proportion of doctoral degree holders compared to California will have estimates that deviate from their actual totals.

- Population diversity: Each state has different demographic compositions. For instance, states with larger metropolitan areas or research universities may have a higher concentration of doctoral degree holders, skewing the estimates when compared to smaller or more rural states.

- Sampling variability: The IPUMS dataset is a sample of the population. Differences between the sample and the actual population (especially in smaller states) can cause the ratio estimator to be inaccurate.


\newpage


# References
